
---
title: "COVID19 API"
author: "Ipsita Datta"
date: "9/28/2021"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
     This document is a vignette to show how to retrieve data from Covid19API.To demonstrate, I'll be interacting with the Covid19 API. I am going to build a few functions to interact with some of the endpoints and explore some of the data I can retrieve.
    
 – The most difficult part of the logic and programming for me is to make a requirement or what I should analysis and how I am going to implement that.
 
 – In,future, First I will sketch a project outline like what are the things is going to discuss or what are the data I should fetch from API.
 
# Requirements 
 To use the functions for interacting with the API, I used the following packages:

- [`tidyverse`](https://www.tidyverse.org/): useful features for data manipulation and visualization
- [`jsonlite`](https://cran.r-project.org/web/packages/jsonlite/): API interaction
- [`httr`](https://cran.r-project.org/web/packages/httr/index.html): For using function to interact with API

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Load in the packages.
library(tidyverse)
library(jsonlite)
library(httr)
```
# API Interaction Functions

Here are the functions to interact with Covid19 API.
## `CountriesName`

I wrote this function to interact with the `CountriesName` endpoint of the Covid19 API.It returns all countries and associated provinces. The country_slug variable is used for country specific data

```{r}
Country <-function(CountryName ="all"){
    #Returns all the available countries and provinces, as well as the country slug for per country requests.
  outputAPI<-GET("https://api.covid19api.com/countries")
  output<- outputAPI$content %>% rawToChar() %>% fromJSON()
  if ( CountryName != "all"){
         #if country name is in country
         if (CountryName %in% output$Country){
         # subset output for just that row.
         output<-output %>% filter(Country == CountryName)
         }
         # if country code in country slug
         else if (CountryName %in% output$Slug ){
          output<-output %>% filter(Slug == CountryName) 
         }
         # if country name is in countrycode/ISO2
         else if (CountryName %in% output$ISO2 ){
          output<-output %>% filter(CountryCode == CountryName) 
         }
  }
    # Return the output data.frame.
  
  return(output)
}

```

## `summary`

This function returns current cases for all the country  and can also return for single country if name passed as argument.
```{r}
summary<-function(CountryName ="all"){
         ## Get all the number of cases from the summary endpoint.
         outputAPI<-fromJSON("https://api.covid19api.com/summary")
         output<-outputAPI$Countries
         if ( CountryName != "all"){
         #if country name is in country
         if (CountryName %in% output$Country){
         # subset output for just that row.
         output<-output %>% filter(Country == CountryName)
         }
         # if country name is in countrycode
         else if (CountryName %in% output$CountryCode ){
          output<-output %>% filter(CountryCode == CountryName) 
         }
         # if country code in country slug
         else if (CountryName %in% output$Slug ){
          output<-output %>% filter(Slug == CountryName) 
         }
          # Otherwise, throw an informative error.
          else {
          message <- paste("ERROR: Argument for country was not found in either","the Country or CountryCode or Slug columns. Try Countries('all') to","find the CountryName you're looking for.")
      stop(message)
          }
}
          # Do nothing if the Country value equals "all".
else {
    
     }
        # Return the output data.frame.
        return(output)
}

```
## `USAdetail`
 Returns all live cases  for All provinces of USA. It can be use for a particular province with particular case type of USA too.These records are pulled every 10 minutes and are ungrouped. Country must be the slug from /countries or /summary. Cases must be one of: Deaths,Confirmed ,Active,Recovered
```{r}
USAdetail<-function(ProvinceName ="all",type  ){
    baseurl<-"https://api.covid19api.com/live/country/united-states"
    output<-fromJSON(baseurl)
    # If Province does not equal "all", check if it is a Province .
        if (ProvinceName != "all" ){
           # If Province is in the Province column, subset output for just that row.
             if (ProvinceName %in% output$Province){
                 output <- output %>%
                           filter(ProvinceName == Province) %>% select(all_of(type),Province)
            }
            # Otherwise, warn the user and return the entire dataframe.
            else {
            message <- paste("WARNING: Argument for Province was not found in either",
                       "the Province  columns. Returning all",
                       "or type should be Deaths,Confirmed ,Active,Recovered.")
           warning(message)
            }
          return(output)
         }
        # Do nothing if the province value equals "all".
        else {# Return the output data.frame.
              return(output)
              }
}
```

## `CountryDatabycase`
 Get List Of Cases Per Country  By Case Type From The First Recorded Case.Returns all cases by case type for a country from the first recorded case. Country must be the country_slug from /countries. Cases must be one of: confirmed, recovered, deaths or Active.

```{r}
CountryDatabycase<-function(Country ="all",Status){
if (Country != "all"){
baseurl<-"https://api.covid19api.com/dayone/country/"
path1<-paste0(baseurl,Country)
endpoints<-"/status/"
fullurl<-paste0(path1,endpoints,Status)
outputAPI <-fromJSON(fullurl)
return(outputAPI)
}
  else{message <- paste("ERROR: Argument for country was not found in  Country ")
      stop(message)
  }
}

```
## ` Allcasesbycountry`
Returns all cases by case type for a country from the first recorded case. Country must be the Slug from /countries or /summary. Cases must be one of: confirmed, recovered, deaths or Active
```{r}
Allcasesbycountry<-function(Country ="all" ,type){
if (Country != "all"){
baseurl<-"https://api.covid19api.com/dayone/country/"
fullurl<-paste0(baseurl,Country)
outputAPI<-fromJSON(fullurl)
output<- outputAPI %>% select(Country,type)
return(output)
}
else{message <- paste("ERROR: Argument for country was not found in  Country ")
      stop(message)
}
}
Allcasesbycountry("Switzerland","Confirmed")
```

## `LiveCount `
Returns all cases by case type for a country from the first recorded case with the latest record being the live count. Country must be the Slug from /countries or /summary. Cases must be one of: confirmed, recovered, deaths
```{r}
Livecount<-function(Country="all",type){
if (Country !="all"){
baseurl<-"https://api.covid19api.com/dayone/country/"
path1<-paste0(baseurl,Country)
fullurl<-paste0(path1,"/status/",type,"/live")
outputAPI<-fromJSON(fullurl)
return(outputAPI)
}
  else{message <- paste("ERROR: Argument for country was not found in  Country ")
      stop(message) }
}
#Livecount("iran","confirmed")
```
# wrrapper function

```{r}
CovidAPI <- function(func, ...){
  ###
  # This function is a wrapper for the other functions. It takes in the name
  # of the function to use as a character and any additional arguments for that
  # function.
  ###
  
  # Find and call the appropriate function using conditional logic.
  if (func == "summary"){
    df <- summary(...)
  }
  else if (func == "Country"){
    df <- Country(...)
  }
   else if (func == "USProvince"){
    df <- USProvince(...)
  }
  else {
    stop("ERROR: Argument for func is not valid!")
  }
  
  # Return the output from the appropriate function.
  return(df)
}
```
# Data Exploration

Now that we can interact with a few of the endpoints of the Covid19 API, let's get some data from them.
First, let's pull A function returns current cases for all the country along with global cases
```{r}
# Get the current cases for all of the country .
currentData<- CovidAPI("summary")
```
Two variables of interest to me are shots per game and shooting percentage. I'm interested in how these two stats relate to a team's win percentage. This variable doesn't exist, so I need to calculate it. I define this as the number of wins divided by the total number of games played. 

```{r}
# Add a column for the death per 10000
currentData <- currentData %>%
  mutate(deathRate = round((Countries.TotalDeaths*10000 )/ Global.TotalDeaths,0),NewconfirmRate=round((Countries.TotalConfirmed*10000 )/ Global.TotalConfirmed,0))

# Create a Bar Plot of deathRate per Country( frist 10 country from the dataset).
plot1<-head(currentData,10) %>% ggplot(aes(  Countries.Country,deathRate,
             fill=deathRate)) +
  geom_col() + 
  #  remove the legend.
  theme(axis.text.x=element_text(angle=90), legend.position="none") +
  # Set the axes labels.
  scale_x_discrete("Country") + 
  scale_y_continuous("DeathRate") +
  # Add a title.
  ggtitle("First 10 country death rate from Current data") 

```
```{r}
# Create a scatter Plot of deathRate per Country( first 10 country from the dataset).
plot2<-currentData %>% ggplot(aes(NewconfirmRate,deathRate,
color=deathRate)) +
  geom_point(size=1, alpha=0.75) + 
  # Rotate the x-axis labls 90 degrees and remove the legend.
  theme( legend.position="none") +
  # Set the axes labels.
  scale_x_continuous("NewconfirmRate") + 
  scale_y_continuous("DeathRate") +
  # Add a title.
  ggtitle("First 10 country death rate from Current data") 
}
```
```{r}
outputAPI<-GET("https://api.covid19api.com/summary")
         data<-fromJSON(rawToChar(outputAPI$content))
         df<-as.data.frame(data)
# Make a box plot of countryname deathrate  by game type.
g<-ggplot(df,aes(Countries.NewConfirmed)) 
  # Add the box plot layer.
 g+geom_boxplot() + 
    # Add a title.
  ggtitle("ConfirmedCases accorss countries") + 
  # Remove the legend because it isn't needed.
  theme(legend.position="none")

```